---
title: "Computing Scores for Wild Data"
output: html_notebook
---

After we obtain the key table which indicates what blocks indicate what score, we then map it into the wild dataset.

We first load the libraries.
```{r, echo=FALSE, include=FALSE}
library(tidyverse)
library(data.table)
```

Then we load the dataset and the key table. For experimenting, we look at the first ten rows of the data

```{r}
project_blocks = read_csv("project_blocks.csv")

```

```{r}
pb_test = project_blocks[1:100,]
pb_test$doifelse
#saveRDS(pb_test,"project1000.rds")
dtWild = filter(pb_test,doifelse > 2)
# dtWild = as.tibble(readRDS("project_blocks.rds"))
dtkey = as.data.table(read_csv("ev_keys.csv"))
dtkey = as.tibble(select(dtkey,block,score,category))
```

Wrangle data to have a cleaned WildClean
```{r}
colnames(dtWild) = gsub("[[:punct:]]","",colnames(dtWild)) #clean column names
dtWild_t = t(dtWild)
block = colnames(dtWild)
dtWildClean= as.tibble(cbind(block,dtWild_t))
rownames(dtWildClean) = c() #remove row names
coln = vector()
for (i in 1:length(rownames(dtWild))){
  coln[i] = paste("respondent",i,sep="")
}
colnames(dtWildClean) = c("block",coln)
```

Testing
```{r}
compute_dataset(dtWildClean,dtkey,1)
```

```{r}
#Helper function used in compute dataset return a boolean to check if the block exists in the data set
#n: threshold of num of blocks to check
  checkBlock = function(dt,blockname,n = 0){
    return(nrow(filter(dt, block==blockname)) != 0 && filter(dt, block==blockname)$num > n)
  }

#Helper function used in compute dataset return a boolean to check if the characters in the block array #exists in the data set
#n: threshold of num of blocks to check
  checkBlockArray = function(dt,blockArray,n = 0){
    for (i in 1:length(blockArray)){
      if (nrow(filter(dt, block==blockArray[i])) != 0 && filter(dt, block==blockArray[i])$num > n){
        return(TRUE)
      }
    }
    return(FALSE)
  }
```

Now lets to generalize the above code
```{r}
#dt = dtWildClean
#dt_key = dtKey, key
#respNum = number of respondent to get from
#PRECONDITION: column titles of scores in dtWildClean must be in the format "respondent#"
compute_dataset = function(dt, dt_key, respNum){
  resp = paste("respondent",respNum,sep="")
  #get scores of respondent #5
  #get dataset of respondent, the blocknames alongside how many exists in the project
  dt_resp = as.tibble(cbind(dt$block,dt[resp]))
  colnames(dt_resp) = c("V1","V2")
  dt_resp$V2 = as.numeric(dt_resp$V2) #set the numbers as numeric from character
  dt_resp_filter = dt_resp %>% #filter out blocks that do not exist
    filter(V2 != 0)
  colnames(dt_resp_filter) = (c("block","num")) 
  
  #map existing blocks with key table
  dt_resp_scores = left_join(dt_resp_filter,dt_key,by="block")
  dt_resp_scores = dt_resp_scores %>% #omit blocks that do not exist
    na.omit()
  
  #RESULT: dt_resp_scores = 
  #scores of respondents, with category, block name, and the number of times it appears
  
  #initialize final data frame
  dt_return = as.tibble(data.frame(category = character(), ev_score = integer()))
  dt_return$category = as.character(dt_return$category)
  
  #compute scores of respondent based on category
  categoryVec = unique(dt_key$category)
  
  for (i in 1:length(categoryVec)){
    category_dt = filter(dt_resp_scores,category == categoryVec[i])
    if (!nrow(category_dt)==0) score_dt = c(categoryVec[i],max(category_dt$score))
    else score_dt = c(categoryVec[i],0)
    dt_return[i,] = score_dt
  }
  
  #assume for variables and sequences
  #for variables, always assume 1, as we assume a sprite is initialized in every project
  if (filter(dt_return,category=="variables")$ev_score == 0) { #if 0, change to 1
    dt_return[4,2] = 1
  }
  #for sequences, assume 1 if green flag block present
  if (filter(dt_return,category=="seq_looping")$ev_score == 0 & checkBlock(dt_resp_scores,"eventhatmorphstartclicked")) {
    dt_return[5,2] = 1
  }
  
  #compute parallelization, initialize location, and initialize looks
  #parallelization = 1 if 2 sprites start at same event, i.e. when two green flag blocks are present
  if(checkBlock(dt_resp_scores,"eventhatmorphstartclicked",1)) {
    dt_return[11,] = c("parallelization",1)
  } else {
    dt_return[11,] = c("parallelization",0)
  }
  
  #if more than one doifelse, assume nested ifelse 
if (checkBlock(dt_resp_scores,"doifelse",1)){
  dt_return[8,2] = 3
}
  
  #initialize location = 1 if green flag block is present(eventhatmorphstartclicked) and setxpos and setypos also exists
  if(checkBlock(dt_resp_scores,"eventhatmorphstartclicked") & (checkBlock(dt_resp_scores,"setxpos") |checkBlock(dt_resp_scores,"setypos"))) {
    dt_return[12,] = c("initialize_location",1)
  } else {
    dt_return[12,] = c("initialize_location",0)
  }
  
  lookBlocks = filter(filter(dt_key,category=="looks"), score != 1)
  #initialize looks = 1 if green flag block is present and looks exist (except say and think)
  if(checkBlock(dt_resp_scores,"eventhatmorphstartclicked") & (checkBlockArray(dt_resp_scores,lookBlocks$block)))   {
    dt_return[13,] = c("initialize_looks",1)
  } else {
    dt_return[13,] = c("initialize_looks",0)
  }
  dt_return
  return(dt_return)
}
#RESULT: dt_return = ev of respondents
```

```{r}
  
```


Now lets use this function to compute scores for all the respondents 
```{r, echo=FALSE, include=FALSE}
for (i in 2:length(colnames(dtWildClean))-1){
  if(i == 1){
    dt_final = compute_dataset(dtWildClean,dtkey,i)
    colnames(dt_final) = c("category",paste("respondent",i,sep=""))
  } else {
  dt_temp = compute_dataset(dtWildClean,dtkey,i)
  colnames(dt_temp) = c("category",paste("respondent",i,sep=""))
  dt_final = left_join(dt_final,dt_temp,by="category")
  }
}
```

```{r}
write.csv(dt_final,"ev_scores.csv")
```

Now we try to map scores for the 5th respondent
```{r}
#get scores of respondent #5
#get dataset of respondent, the blocknames alongside how many exists in the project
dt_resp = as.tibble(cbind(dtWildClean$block,dtWildClean["respondent5"]))
colnames(dt_resp) = c("V1","V2")
dt_resp$V2 = as.numeric(dt_resp$V2) #set the numbers as numeric from character
dt_resp_filter = dt_resp %>% #filter out blocks that do not exist
  filter(V2 != 0)
colnames(dt_resp_filter) = (c("block","num")) 
dt_resp_scores = left_join(dt_resp_filter,dtkey,by="block") #map existing blocks with key table
dt_resp_scores = dt_resp_scores %>% #omit blocks that do not exist
  na.omit()

#RESULT: dt_resp_scores = 
#scores of respondents, with category, block name, and the number of times it appears

#initialize final data frame
dt_return = as.tibble(data.frame(category = character(), ev_score = integer()))
dt_return$category = as.character(dt_return$category)

dt_resp_scores_ordered = dt_resp_scores[order(dt_resp_scores$score),] #order the scores for view

#compute scores of respondent
categoryVec = unique(dtkey$category)

for (i in 1:length(categoryVec)){
  category_dt = filter(dt_resp_scores,category == categoryVec[i])
  if (!nrow(category_dt)==0) score_dt = c(categoryVec[i],max(category_dt$score))
  else score_dt = c(categoryVec[i],0)
  dt_return[i,] = score_dt
}

#return a boolean to check if the block exists in the data set
#n: how many blocks to check
checkBlock = function(dt,blockname,n = 0){
  return(nrow(filter(dt_resp_scores, block==blockname)) != 0 && filter(dt_resp_scores, block==blockname)$num > n)
}

checkBlockArray = function(dt,blockArray,n = 0){
  for (i in 1:length(blockArray)){
    if (nrow(filter(dt_resp_scores, block==blockArray[i])) != 0 && filter(dt_resp_scores, block==blockArray[i])$num > n){
      return(TRUE)
    }
  }
  return(FALSE)
}

#assume for variables and sequences
#for variables, always assume 1, as we assume a sprite is initialized in every project
if (filter(dt_return,category=="variables")$ev_score == 0) { #if 0, change to 1
  dt_return[4,2] = 1
}
#for sequences, assume 1 if green flag block present
if (filter(dt_return,category=="seq_looping")$ev_score == 0 & checkBlock(dt_resp_scores,"eventhatmorphstartclicked")) {
  dt_return[5,2] = 1
}



#compute parallelization, initialize location, and initialize looks
#parallelization = 1 if 2 sprites start at same event, i.e. when two green flag blocks are present
if(checkBlock(dt_resp_scores,"eventhatmorphstartclicked",1)) {
  dt_return[11,] = c("parallelization",1)
} else {
  dt_return[11,] = c("parallelization",0)
}

#initialize location = 1 if green flag block is present(eventhatmorphstartclicked) and setxpos and setypos also exists
if(checkBlock(dt_resp_scores,"eventhatmorphstartclicked") & (checkBlock(dt_resp_scores,"setxpos") |checkBlock(dt_resp_scores,"setypos"))) {
  dt_return[12,] = c("initialize_location",1)
} else {
  dt_return[12,] = c("initialize_location",0)
}


lookBlocks = filter(filter(dtkey,category=="looks"), score != 1)
#initialize looks = 1 if green flag block is present and looks exist (except say and think)
if(checkBlock(dt_resp_scores,"eventhatmorphstartclicked") & (checkBlockArray(dt_resp_scores,lookBlocks$block))) {
  dt_return[13,] = c("initialize_looks",1)
} else {
  dt_return[13,] = c("initialize_looks",0)
}

#RESULT: dt_return = ev of respondents
dt_return
dt_resp_scores_ordered
```
```{r}
df = readRDS("project_blocks.rds")
```